---
execute:
  warning: false
---

# Invariance

We examined whether our scale's psychometric properties were equal across groups with and without insomnia symptoms and across time points, comparing baseline test scores to a second administration taken 14 days later. To do this, we tested invariance by restricting the measurement properties across groups/time in increasing levels of stringency. We tested in a sequence of factor structure (configural invariance), factor loadings (metric invariance), items' intercepts (scalar invariance), and items' residual variances (strict invariance) [@flake2021]. If a level of non-invariance was detected, no further tests were conducted. The criteria used to evaluate model fit were the chi-squared difference test for goodness of fit and alternative fit indices like $\Delta$CFI, $\Delta$RMSEA, and $\Delta$SRMR

## Loading the required packages

```{r}
library(semTools)
library(kableExtra)
```


```{r, include=FALSE}

mydata <- readr::read_csv("data/clean_data/clean_data.csv")

arm1 <- mydata |> 
  dplyr::filter(redcap_event_name == "elegibilidade_arm_1",
                !dplyr::if_all(dplyr::starts_with("sas_"), ~ is.na(.))) |> 
  dplyr::mutate(group = dplyr::case_when(
    dsm_1 == 0 &
    dsm_2 == 0 &
    dsm_3 == 0 &
    dsm_4 == 0 &
    dsm_5 == 0 &
    isi_total < 8 ~ "good_sleepers",
    TRUE ~ "bad_sleepers")) |> 
  dplyr::select(record_id, group, paste0("sas_", 1:6))

items <- arm1[,2:8]
```


## Longitudinal invariance

```{r longitudinal-data, echo=FALSE}
data_longitudinal <- mydata |> 
  dplyr::select(record_id, paste0("sas_", 1:6), redcap_event_name) |> 
  dplyr::filter(!dplyr::if_all(dplyr::starts_with("sas_"), ~ is.na(.))) |> 
  tidyr::pivot_longer(cols = sas_1:sas_6, names_to = "item", values_to = "score") |> 
  dplyr::mutate(item = ifelse(redcap_event_name == "elegibilidade_arm_1", paste0(item, ".1"), paste0(item, ".2"))) |> 
  dplyr::select(-redcap_event_name) |> 
  tidyr::pivot_wider(names_from = item, values_from = score)
```

### Configural invariance

```{r configural-model}

long_configural_model <- '

SAS.1 =~ sas_1.1 + sas_2.1 + sas_3.1 + sas_4.1 + sas_5.1 + sas_6.1

SAS.2 =~ sas_1.2 + sas_2.2 + sas_3.2 + sas_4.2 + sas_5.2 + sas_6.2

'

longFacNames <- list (SAS = c("SAS.1", "SAS.2"))

syntax_long_config <- measEq.syntax(configural.model = long_configural_model,
                               data = data_longitudinal,
                               parameterization = "theta",
                               ID.fac = "std.lv", ID.cat = "Wu.Estabrook.2016",
                               longFacNames = longFacNames)

mod_long_config <- as.character(syntax_long_config)

fit_long_config <- cfa(mod_long_config, 
                      data = data_longitudinal, 
                      estimator = "DWLS", 
                      std.lv = TRUE)

semTools::fitmeasures(fit_long_config, fit.measures = c("rmsea", "rmsea.ci.lower", "rmsea.ci.upper", "cfi", "srmr_bentler")) |> 
  kbl(col.names = c("Index", "Value"))
```

### Metric invariance

```{r metric-invariance}
syntax_long_metric <- measEq.syntax(configural.model = long_configural_model,
                               data = data_longitudinal,
                               parameterization = "theta",
                               ID.fac = "std.lv", ID.cat = "Wu.Estabrook.2016",
                               longFacNames = longFacNames,
                               long.equal  = c("thresholds","loadings"))

mod_long_metric <- as.character(syntax_long_metric)

fit_long_metric <- cfa(mod_long_metric, data = data_longitudinal, estimator = "DWLS",
                  parameterization = "theta", std.lv = TRUE)

conf_met_long <- compareFit(fit_long_config, fit_long_metric)

lavTestLRT(fit_long_config, fit_long_metric, method="satorra.bentler.2010") |> 
  broom::tidy() |> 
  kbl()
```

### Scalar invariance

```{r scalar-invariance}
syntax_long_scalar <- measEq.syntax(configural.model = long_configural_model,
                               data = data_longitudinal,
                               parameterization = "theta",
                               ID.fac = "std.lv", ID.cat = "Wu.Estabrook.2016",
                               longFacNames = longFacNames,
                               long.equal  = c("thresholds","loadings",
                                               "intercepts"))
mod_long_scalar <- as.character(syntax_long_scalar)
fit_long_scalar <- cfa(mod_long_scalar, data = data_longitudinal, estimator = "DWLS",
                  parameterization = "theta", std.lv = TRUE)

anova(fit_long_config, fit_long_metric, fit_long_scalar, method = "satorra.bentler.2010") |> 
  broom::tidy() |> 
  kbl()
```

### Strict invariance

```{r}
syntax_long_strict <- measEq.syntax(configural.model = long_configural_model,
                               data = data_longitudinal,
                               parameterization = "theta",
                               ID.fac = "std.lv", ID.cat = "Wu.Estabrook.2016",
                               longFacNames = longFacNames,
                               long.equal  = c("thresholds","loadings",
                                               "intercepts", "residuals"))
mod_long_strict <- as.character(syntax_long_strict)
fit_long_strict <- cfa(mod_long_strict, data = data_longitudinal, estimator = "DWLS",
                  parameterization = "theta", std.lv = TRUE)

anova(fit_long_config, fit_long_metric, fit_long_scalar, fit_long_strict, method = "satorra.bentler.2010")|> 
  broom::tidy() |> 
  kbl()
```

### Longitudinal fit differences

```{r longitudinal-fit-differences}
inv_long <- anova(fit_long_config, fit_long_metric, fit_long_scalar, fit_long_strict,
                  method = "satorra.bentler.2010") |> 
  tibble::as.tibble()

fit_stats_strict <- fitmeasures(fit_long_strict, 
                                fit.measures = c("rmsea", "rmsea.ci.lower", "rmsea.ci.upper", "cfi", "srmr_bentler"))

fit_stats_scalar <- fitmeasures(fit_long_scalar, 
                                fit.measures = c("rmsea", "rmsea.ci.lower", "rmsea.ci.upper", "cfi", "srmr_bentler"))

fit_stats_metric <- fitmeasures(fit_long_metric, 
                                fit.measures = c("rmsea", "rmsea.ci.lower", "rmsea.ci.upper", "cfi", "srmr_bentler"))

fit_stats_configural <- fitmeasures(fit_long_config, 
                                fit.measures = c("rmsea", "rmsea.ci.lower", "rmsea.ci.upper", "cfi", "srmr_bentler"))

long_fit_diffs <- tibble::enframe(c(
        fit_stats_configural - fit_stats_metric,
        fit_stats_metric - fit_stats_scalar,
        fit_stats_scalar - fit_stats_strict)) |> 
  dplyr::mutate(model = rep(c("metric", "scalar", "strict"), each=5),
                .before = name) |> 
  tidyr::pivot_wider(names_from = "name", values_from = "value") 

long_fit_diffs |> 
  kableExtra::kbl()
```

## Invariance across groups (good x bad sleepers)

```{r}
model_sas <- "
SAS =~ sas_1 + sas_2 + sas_3 + sas_4 + sas_5 + sas_6
"
```

### Configural invariance

```{r configural-model-group}
syntax.config.group <- measEq.syntax(configural.model = model_sas,
                               data = items,
                               group = "group",
                               parameterization = "theta",
                               ID.fac = "std.lv", ID.cat = "Wu.Estabrook.2016")
mod.config.group <- as.character(syntax.config.group)

configural.fit.group <- cfa(mod.config.group, 
                           group = "group",
                           data = items,
                           estimator = "DWLS",
                           std.lv = TRUE)

semTools::fitmeasures(configural.fit.group, fit.measures = c("rmsea", "rmsea.ci.lower", "rmsea.ci.upper", "cfi", "srmr_bentler")) |> 
  kbl(col.names = c("Index", "Value"))
```

### Metric invariance

```{r metric-invariance-group}
syntax.metric.group <- measEq.syntax(configural.model = model_sas,
                               data = items,
                               group = "group",
                               parameterization = "theta",
                               ID.fac = "std.lv", ID.cat = "Wu.Estabrook.2016",
                               group.equal  = c("loadings"))

mod.metric.group <- as.character(syntax.metric.group)

fit.metric.group <- cfa(mod.metric.group, group = "group",
                           data = items,
                           estimator = "DWLS",
                           std.lv = TRUE)

conf.met <- compareFit(configural.fit.group, fit.metric.group, argsLRT = list(method = "satorra.bentler.2010"))


lavTestLRT(configural.fit.group, fit.metric.group, method="satorra.bentler.2010") |> 
  broom::tidy() |> 
  kbl()
```

```{r inspecting-metric-inv}

lavaan::modindices(fit.metric.group) |> 
  dplyr::arrange(-mi) |> 
  head(10) |> 
  kbl()
```
